---
title: "Week 11 Lab: Skill Wrap-Up"
subtitle: "AAI1001 – Data Engineering and Visualization | AY25/26 Tri 2 | Assessment 5 (Teamwork 2)"
date: today
date-format: "dddd MMM DD, YYYY"
author:
  - name: "Goh Jason Fadil (2502643)"
    affiliations:
      - name: "Singapore Institute of Technology"
  - name: "Venus Low (2503741)"
  - name: "Choong Wan Qin (2402311)"
  - name: "Chi Wei Seng (2401493)"
  - name: "Koh Soon Chuan (2400951)"
  
format:
  html:
    embed-resources: true
    number-sections: true
    toc: true
    toc-depth: 3
    toc-title: "Table of Contents"
    toc-location: left
    theme: flatly
    highlight-style: github
    code-fold: false
    df-print: paged
execute:
  echo: true
  warning: false
  message: false
---
# Analyzing and Visualizing DJIA Data

## Import and Clean Data
```{r}
library(tidyverse)
library(readxl)

monthly_dj <- read_excel("dja-performance-report-monthly.xlsx", skip = 5) |>
  select(`Effective Date`, `Close Value`) |>
  rename(date = `Effective Date`, close_value = `Close Value`) |>
  mutate(
    date = as.Date(date, format = "%m/%d/%Y"),
    close_value = as.numeric(close_value)
  ) |>
  drop_na() |>
  arrange(date)

monthly_dj |> slice_head(n = 10)
```
## Identify Gap Months
```{r}
gap_months <- monthly_dj |>
  mutate(
    month = month(date),
    year = year(date),
    months_to_next_row = interval(date, lead(date)) %/% months(1)
  ) |>
  filter(months_to_next_row > 1) |>
  select(date, close_value, month, year, months_to_next_row)

gap_months
```

The gap occurs from August 1914 to December 1914, 5 missing months. This coincides with the closure of the New York Stock Exchange at the outbreak of World War I, as European nations began liquidating US securities, threatening a financial panic. The NYSE suspended trading from July 31 to December 12, 1914.

## Visualizing DJIA Performance by Month
```{r}
monthly_dj <- monthly_dj |>
  mutate(
    rel_change_pct = (close_value / lag(close_value) - 1) * 100,
    month_abb = factor(month.abb[month(date)], levels = month.abb)
  )
```

### Visualization A - Violin and Box Plots
```{r}
median_line <- monthly_dj |>
  drop_na(rel_change_pct) |>
  group_by(month_abb) |>
  summarise(median_val = median(rel_change_pct))

monthly_dj |>
  drop_na(rel_change_pct) |>
  ggplot(aes(x = month_abb, y = rel_change_pct)) +
  geom_violin(fill = "steelblue", alpha = 0.4, trim = FALSE) +
  geom_boxplot(width = 0.2, outlier.shape = NA, coef = 0) +
  geom_line(
    data = median_line,
    aes(x = month_abb, y = median_val, group = 1),
    colour = "firebrick", linewidth = 0.8
  ) +
  geom_point(
    data = median_line,
    aes(x = month_abb, y = median_val),
    colour = "firebrick", size = 2
  ) +
  labs(
    title = "DJIA Monthly Returns — Violin + Box Plot",
    x = "Month", y = "Relative Change (%)"
  ) +
  theme_minimal()
```

### Visualization B - Faceted Frequency Polygons 
```{r}
mean_line <- monthly_dj |>
  drop_na(rel_change_pct) |>
  group_by(month_abb) |>
  summarise(mean_val = mean(rel_change_pct))

monthly_dj |>
  drop_na(rel_change_pct) |>
  ggplot(aes(x = rel_change_pct)) +
  geom_freqpoly(binwidth = 2, colour = "steelblue") +
  geom_rug(alpha = 0.4) +
  geom_vline(
    data = mean_line,
    aes(xintercept = mean_val),
    linetype = "dashed", colour = "firebrick"
  ) +
  facet_wrap(~ month_abb, nrow = 4, ncol = 3) +
  labs(
    title = "DJIA Monthly Returns — Frequency Polygons by Month",
    x = "Relative Change (%)", y = "Count"
  ) +
  theme_minimal()
```


