
```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(tidyquant)
library(gt)
```
```{r}
#| label: task-2-2-1

ticker_btc <- "BTC-USD"
ticker_gold <- "GC=F"
start_date <- "2015-01-01"
end_date <- "2026-02-28"

monthly_btc <- tq_get(ticker_btc, from = start_date, to = end_date) |>
  drop_na(adjusted) |>
  tq_transmute(select = adjusted, mutate_fun = to.period, period = "months")

monthly_gold <- tq_get(ticker_gold, from = start_date, to = end_date) |>
  drop_na(adjusted) |>
  tq_transmute(select = adjusted, mutate_fun = to.period, period = "months")

head(monthly_btc, 10)
head(monthly_gold, 10)
```

```{r}
#| label: task-2-2-dj

monthly_dj <- tq_get("^DJI", from = "2015-01-01", to = "2026-02-28") |>
  drop_na(adjusted) |>
  tq_transmute(select = adjusted, mutate_fun = to.period, period = "months") |>
  rename(date = date, close_value = adjusted)
```


```{r}
#| label: task-2-2-2
#| fig-width: 10
#| fig-height: 6

# Standardise each asset into same 3-column format
dj_plot <- monthly_dj |>
  select(month = date, close_value) |>
  mutate(asset = "DJIA")

btc_plot <- monthly_btc |>
  select(month = date, close_value = adjusted) |>
  mutate(asset = "BTC")

gold_plot <- monthly_gold |>
  select(month = date, close_value = adjusted) |>
  mutate(asset = "Gold")

# Stack all 3 into one long data frame
combined <- bind_rows(dj_plot, btc_plot, gold_plot) |>
  mutate(asset = factor(asset, levels = c("DJIA", "BTC", "Gold")))

ggplot(combined, aes(x = month, y = close_value, color = asset)) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~asset, scales = "free_y", ncol = 1) +
  scale_color_manual(values = c("DJIA" = "steelblue", 
                                 "BTC" = "orange", 
                                 "Gold" = "gold3")) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  labs(
    title = "Monthly Closing Prices: DJIA, BTC, and Gold (2015–2026)",
    x = "Date",
    y = "Price (USD)",
    color = "Asset"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

```{r}
#| label: task-2-2-3

compute_drawdown <- function(prices) {
  drawdowns <- numeric(length(prices))
  peak <- prices[1]
  for (i in seq_along(prices)) {
    if (prices[i] > peak) peak <- prices[i]
    drawdowns[i] <- (prices[i] - peak) / peak * 100
  }
  return(drawdowns)
}

drawdowns_gold <- compute_drawdown(monthly_gold$adjusted)
drawdowns_dj   <- compute_drawdown(monthly_dj$close_value)
drawdowns_btc  <- compute_drawdown(monthly_btc$adjusted)

biggest_drawdown_gold <- min(drawdowns_gold)
biggest_drawdown_dj   <- min(drawdowns_dj)
biggest_drawdown_btc  <- min(drawdowns_btc)

biggest_drawdown_gold
biggest_drawdown_dj
biggest_drawdown_btc
```
The drawdown analysis confirms our hypothesis that volatility increases from Gold to DJIA to BTC. Gold recorded the smallest largest drawdown at approximately -17.07%, reflecting its reputation as a stable store of value over centuries. The DJIA showed a moderate drawdown of -23.20%, consistent with the broader stock market's resilience over time. BTC, however, experienced the largest drawdown at approximately -75.57%, validating its classification as a highly volatile speculative asset. These results support the conclusion that volatility increases progressively from Gold to DJIA to BTC.



