---
title: "Team 30 - Assignment 2 Teamwork"
author: "Goh Jason Fadil (2502643), Venus Low (2503741), Choong Wan Qin (2402311), Chi Wei Seng (2401493), Koh Soon Chuan (2400951)"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

# Task 1: English Premier League Table 2019–2020

This section of the document contains the **English Premier League 2019/20 Final Standings** from raw match results using `tidyverse` for data wrangling and `gt` for table styling.

---

### Load Libraries

```{r}
library(tidyverse)
library(gt)
```

---

### Load Data

```{r}
results_raw <- read_csv("epl_results.csv")
teams_df    <- read_csv("epl_teams.csv")

teams <- teams_df$abbr
```

---

### Parse Match Results

Each match is parsed into two rows — one per team — recording goals scored, goals conceded, and points earned.

```{r}
records <- list()

for (i in seq_len(nrow(results_raw))) {
  home <- results_raw$home_team[i]

  for (away in teams) {
    if (away == home) next

    score <- results_raw[[away]][i]
    if (is.na(score) || score == "-") next

    parts <- str_split(score, "-")[[1]]
    hg    <- as.integer(parts[1])
    ag    <- as.integer(parts[2])

    # Home team row
    records[[length(records) + 1]] <- tibble(
      team          = home,
      goals_for     = hg,
      goals_against = ag,
      pts           = case_when(hg > ag ~ 3L, hg == ag ~ 1L, TRUE ~ 0L)
    )

    # Away team row
    records[[length(records) + 1]] <- tibble(
      team          = away,
      goals_for     = ag,
      goals_against = hg,
      pts           = case_when(ag > hg ~ 3L, ag == hg ~ 1L, TRUE ~ 0L)
    )
  }
}

matches <- bind_rows(records)
```

---

### Aggregate & Rank Teams

Teams are ranked by **points**, then **goal difference**, then **goals scored** as tiebreakers.

```{r}
league <- matches |>
  group_by(team) |>
  summarise(
    points        = sum(pts),
    goals_scored  = sum(goals_for),
    goals_against = sum(goals_against),
    .groups       = "drop"
  ) |>
  mutate(goal_diff = goals_scored - goals_against) |>
  arrange(desc(points), desc(goal_diff), desc(goals_scored)) |>
  mutate(rank = row_number()) |>
  left_join(teams_df |> rename(team_full = name), by = c("team" = "abbr")) |>
  select(rank, team = team_full, points, goal_diff, goals_scored)
```

---

### Generate League Table

Top 4 qualify for the **UEFA Champions League** (highlighted in blue); bottom 3 are **relegated** (highlighted in red).

```{r}
league |>
  gt() |>

  # Title & subtitle
  tab_header(
    title    = md("**\u26bd Premier League 2019/20 \u2014 Final Standings**"),
    subtitle = md("*Points \u00b7 Goal Difference \u00b7 Goals Scored*")
  ) |>

  # Column labels
  cols_label(
    rank         = md("**#**"),
    team         = md("**Club**"),
    points       = md("**Pts**"),
    goal_diff    = md("**GD**"),
    goals_scored = md("**GF**")
  ) |>

  # Alignment
  cols_align(align = "center", columns = c(rank, points, goal_diff, goals_scored)) |>
  cols_align(align = "left",   columns = team) |>

  # Row striping
  opt_row_striping() |>

  # Font
  opt_table_font(font = list(google_font("Roboto"), default_fonts())) |>

  # Global table options
  tab_options(
    heading.background.color       = "#38003c",
    heading.title.font.size        = px(20),
    heading.subtitle.font.size     = px(13),
    column_labels.background.color = "#00ff85",
    column_labels.font.weight      = "bold",
    column_labels.font.size        = px(13),
    row.striping.background_color  = "#f9f9f9",
    table.border.top.color         = "#38003c",
    table.border.bottom.color      = "#38003c",
    table_body.hlines.color        = "#e0e0e0",
    table.font.size                = px(13),
    data_row.padding               = px(6)
  ) |>

  # White text on dark purple heading
  tab_style(
    style     = list(cell_text(color = "white", weight = "bold")),
    locations = cells_title(groups = c("title", "subtitle"))
  ) |>

  # Top 4 — Champions League — blue
  tab_style(
    style     = list(cell_fill(color = "#cce5ff"), cell_text(weight = "bold")),
    locations = cells_body(rows = rank <= 4)
  ) |>

  # Bottom 3 — Relegation — red
  tab_style(
    style     = list(cell_fill(color = "#ffd6d6"), cell_text(color = "#cc0000")),
    locations = cells_body(rows = rank >= 18)
  ) |>

  # Bold rank column
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_body(columns = rank)
  ) |>

  # Positive GD in green
  tab_style(
    style     = cell_text(color = "#007a33", weight = "bold"),
    locations = cells_body(columns = goal_diff, rows = goal_diff > 0)
  ) |>

  # Negative GD in red
  tab_style(
    style     = cell_text(color = "#cc0000", weight = "bold"),
    locations = cells_body(columns = goal_diff, rows = goal_diff < 0)
  ) |>

  # Source note legend
  tab_source_note(
    source_note = md(
      "\U0001f535 Top 4: **UEFA Champions League** &nbsp;|&nbsp; \U0001f534 Bottom 3: **Relegated**"
    )
  ) |>

  # Champion footnote
  tab_footnote(
    footnote  = md("\U0001f3c6 **League Champions**"),
    locations = cells_body(columns = team, rows = rank == 1)
  )
```

---

# Task 2: Investment Market Price Analysis

---

# Task 3: Visualizing Singapore Tourists Spending 

---

# Task 4: Reflection on the Use of GenAI

---
