---
title: "Week 11 Lab: Skill Wrap-Up"
subtitle: "AAI1001 – Data Engineering and Visualization | AY25/26 Tri 2 | Assessment 5 (Teamwork 2)"
date: today
date-format: "dddd MMM DD, YYYY"
author:
  - name: "Goh Jason Fadil (2502643)"
    affiliations:
      - name: "Singapore Institute of Technology"
  - name: "Venus Low (2503741)"
    affiliations:
      - name: "Singapore Institute of Technology"
  - name: "Choong Wan Qin (2402311)"
    affiliations:
      - name: "Singapore Institute of Technology"
  - name: "Chi Wei Seng (2401493)"
    affiliations:
      - name: "Singapore Institute of Technology"
  - name: "Koh Soon Chuan (2400951)"
    affiliations:
      - name: "Singapore Institute of Technology"

format:
  html:
    embed-resources: true
    number-sections: true
    toc: true
    toc-depth: 3
    toc-title: "Table of Contents"
    toc-location: left
    theme: flatly
    highlight-style: github
    code-fold: false
    df-print: paged
execute:
  echo: true
  warning: false
  message: false
---

# Task 1: English Premier League Table 2019–2020

This section of the document contains the **English Premier League 2019/20 Final Standings** from raw match results using `tidyverse` for data wrangling and `gt` for table styling.

---

## Load Libraries

```{r}
library(tidyverse)
library(gt)
```

---

## Load Data

```{r}
results_raw <- read_csv("epl_results.csv")
teams_df    <- read_csv("epl_teams.csv")

teams <- teams_df$abbr
```

---

## Parse Match Results

Each match is parsed into two rows — one per team — recording goals scored, goals conceded, and points earned.

```{r}
records <- list()

for (i in seq_len(nrow(results_raw))) {
  home <- results_raw$home_team[i]

  for (away in teams) {
    if (away == home) next

    score <- results_raw[[away]][i]
    if (is.na(score) || score == "-") next

    parts <- str_split(score, "-")[[1]]
    hg    <- as.integer(parts[1])
    ag    <- as.integer(parts[2])

    # Home team row
    records[[length(records) + 1]] <- tibble(
      team          = home,
      goals_for     = hg,
      goals_against = ag,
      pts           = case_when(hg > ag ~ 3L, hg == ag ~ 1L, TRUE ~ 0L)
    )

    # Away team row
    records[[length(records) + 1]] <- tibble(
      team          = away,
      goals_for     = ag,
      goals_against = hg,
      pts           = case_when(ag > hg ~ 3L, ag == hg ~ 1L, TRUE ~ 0L)
    )
  }
}

matches <- bind_rows(records)
```

---

## Aggregate & Rank Teams

Teams are ranked by **points**, then **goal difference**, then **goals scored** as tiebreakers.

```{r}
league <- matches |>
  group_by(team) |>
  summarise(
    points        = sum(pts),
    goals_scored  = sum(goals_for),
    goals_against = sum(goals_against),
    .groups       = "drop"
  ) |>
  mutate(goal_diff = goals_scored - goals_against) |>
  arrange(desc(points), desc(goal_diff), desc(goals_scored)) |>
  mutate(rank = row_number()) |>
  left_join(teams_df |> rename(team_full = name), by = c("team" = "abbr")) |>
  select(rank, team = team_full, points, goal_diff, goals_scored)
```

---

## Generate League Table

Top 4 qualify for the **UEFA Champions League** (highlighted in blue); bottom 3 are **relegated** (highlighted in red).

```{r}
league |>
  gt() |>

  # Title & subtitle
  tab_header(
    title    = md("**\u26bd Premier League 2019/20 \u2014 Final Standings**"),
    subtitle = md("*Points \u00b7 Goal Difference \u00b7 Goals Scored*")
  ) |>

  # Column labels
  cols_label(
    rank         = md("**#**"),
    team         = md("**Club**"),
    points       = md("**Pts**"),
    goal_diff    = md("**GD**"),
    goals_scored = md("**GF**")
  ) |>

  # Alignment
  cols_align(align = "center", columns = c(rank, points, goal_diff, goals_scored)) |>
  cols_align(align = "left",   columns = team) |>

  # Row striping
  opt_row_striping() |>

  # Font
  opt_table_font(font = list(google_font("Roboto"), default_fonts())) |>

  # Global table options
  tab_options(
    heading.background.color       = "#38003c",
    heading.title.font.size        = px(20),
    heading.subtitle.font.size     = px(13),
    column_labels.background.color = "#00ff85",
    column_labels.font.weight      = "bold",
    column_labels.font.size        = px(13),
    row.striping.background_color  = "#f9f9f9",
    table.border.top.color         = "#38003c",
    table.border.bottom.color      = "#38003c",
    table_body.hlines.color        = "#e0e0e0",
    table.font.size                = px(13),
    data_row.padding               = px(6)
  ) |>

  # White text on dark purple heading
  tab_style(
    style     = list(cell_text(color = "white", weight = "bold")),
    locations = cells_title(groups = c("title", "subtitle"))
  ) |>

  # Top 4 — Champions League — blue
  tab_style(
    style     = list(cell_fill(color = "#cce5ff"), cell_text(weight = "bold")),
    locations = cells_body(rows = rank <= 4)
  ) |>

  # Bottom 3 — Relegation — red
  tab_style(
    style     = list(cell_fill(color = "#ffd6d6"), cell_text(color = "#cc0000")),
    locations = cells_body(rows = rank >= 18)
  ) |>

  # Bold rank column
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_body(columns = rank)
  ) |>

  # Positive GD in green
  tab_style(
    style     = cell_text(color = "#007a33", weight = "bold"),
    locations = cells_body(columns = goal_diff, rows = goal_diff > 0)
  ) |>

  # Negative GD in red
  tab_style(
    style     = cell_text(color = "#cc0000", weight = "bold"),
    locations = cells_body(columns = goal_diff, rows = goal_diff < 0)
  ) |>

  # Source note legend
  tab_source_note(
    source_note = md(
      "\U0001f535 Top 4: **UEFA Champions League** &nbsp;|&nbsp; \U0001f534 Bottom 3: **Relegated**"
    )
  ) |>

  # Champion footnote
  tab_footnote(
    footnote  = md("\U0001f3c6 **League Champions**"),
    locations = cells_body(columns = team, rows = rank == 1)
  )
```

---

# Task 2: Investment Market Price Analysis

## Analyzing and Visualizing DJIA Data

::: {.callout-note}
**Dataset:** Dow Jones Industrial Average (DJIA) monthly closing values from May 1896 to present, sourced from the Bloomberg DJIA Performance Report.
:::

This section covers the import, cleaning, and exploratory analysis of the Dow Jones Industrial Average (DJIA) monthly performance data sourced from Bloomberg.

### Import and Clean Data

The raw Excel file contains 5 header rows of metadata before the actual data begins. We skip these and extract only the two relevant columns — `Effective Date` and `Close Value` — coercing them to the appropriate R types.

```{r}
#| label: import-data

library(tidyverse)
library(readxl)

monthly_dj <- read_excel("dja-performance-report-monthly.xlsx", skip = 5) |>
  select(`Effective Date`, `Close Value`) |>
  rename(date = `Effective Date`, close_value = `Close Value`) |>
  mutate(
    date = as.Date(date, format = "%m/%d/%Y"),
    close_value = as.numeric(close_value)
  ) |>
  drop_na() |>
  arrange(date)

monthly_dj |> slice_head(n = 10)
```

### Identify Gap Months

We identify months where the gap to the next recorded date exceeds one month, indicating missing data in the series. The `months_to_next_row` column captures the number of calendar months between each row and the one that follows it.

```{r}
#| label: gap-months

gap_months <- monthly_dj |>
  mutate(
    month = month(date),
    year = year(date),
    months_to_next_row = interval(date, lead(date)) %/% months(1)
  ) |>
  filter(months_to_next_row > 1) |>
  select(date, close_value, month, year, months_to_next_row)

gap_months
```

::: {.callout-important}
**Historical Context:** The gap spans August to December 1914 — five missing months. This coincides with the New York Stock Exchange closure at the outbreak of World War I, as European nations liquidated US securities, threatening financial panic. The NYSE suspended trading from July 31 to December 12, 1914.
:::

### Visualizing DJIA Performance by Month

Monthly relative returns are computed and visualised using two complementary approaches to reveal distributional patterns across the calendar year. The `rel_change_pct` column captures the month-on-month percentage change in closing value, and `month_abb` provides an ordered factor for correct calendar ordering in plots.

```{r}
#| label: append-columns

monthly_dj <- monthly_dj |>
  mutate(
    rel_change_pct = (close_value / lag(close_value) - 1) * 100,
    month_abb = factor(month.abb[month(date)], levels = month.abb)
  )
```

#### Visualization A — Violin and Box Plots

The violin plot shows the full distribution of monthly returns, while the overlaid box plot highlights the interquartile range and median. The red line connects successive monthly medians to reveal any seasonal trends across the year.

```{r}
#| label: vis-a
#| fig-width: 10
#| fig-height: 6

median_line <- monthly_dj |>
  drop_na(rel_change_pct) |>
  group_by(month_abb) |>
  summarise(median_val = median(rel_change_pct))

monthly_dj |>
  drop_na(rel_change_pct) |>
  ggplot(aes(x = month_abb, y = rel_change_pct)) +
  geom_violin(fill = "steelblue", alpha = 0.4, trim = FALSE) +
  geom_boxplot(width = 0.2, outlier.shape = NA, coef = 0) +
  geom_line(
    data = median_line,
    aes(x = month_abb, y = median_val, group = 1),
    colour = "firebrick", linewidth = 0.8
  ) +
  geom_point(
    data = median_line,
    aes(x = month_abb, y = median_val),
    colour = "firebrick", size = 2
  ) +
  labs(
    title = "DJIA Monthly Returns — Violin and Box Plots",
    subtitle = "Red line connects successive monthly medians",
    caption = "Source: Bloomberg DJIA Performance Report",
    x = "Month",
    y = "Relative Change (%)"
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"))
```

#### Visualization B — Faceted Frequency Polygons

Each facet shows the distribution of monthly returns as a frequency polygon with a rug plot along the x-axis indicating individual observations. The dashed red vertical line marks the mean return for each month.

```{r}
#| label: vis-b
#| fig-width: 10
#| fig-height: 10

mean_line <- monthly_dj |>
  drop_na(rel_change_pct) |>
  group_by(month_abb) |>
  summarise(mean_val = mean(rel_change_pct))

monthly_dj |>
  drop_na(rel_change_pct) |>
  ggplot(aes(x = rel_change_pct)) +
  geom_freqpoly(binwidth = 2, colour = "steelblue") +
  geom_rug(alpha = 0.4) +
  geom_vline(
    data = mean_line,
    aes(xintercept = mean_val),
    linetype = "dashed",
    colour = "firebrick"
  ) +
  facet_wrap(~ month_abb, nrow = 4, ncol = 3) +
  labs(
    title = "DJIA Monthly Returns — Frequency Polygons by Month",
    subtitle = "Dashed red line indicates the mean return per month",
    caption = "Source: Bloomberg DJIA Performance Report",
    x = "Relative Change (%)",
    y = "Count"
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"))
```

## Comparing DJIA with Bitcoin and Gold Prices
```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(tidyquant)
library(gt)
```
```{r}
#| label: task-2-2-1

ticker_btc <- "BTC-USD"
ticker_gold <- "GC=F"
start_date <- "2015-01-01"
end_date <- "2026-02-28"

monthly_btc <- tq_get(ticker_btc, from = start_date, to = end_date) |>
  drop_na(adjusted) |>
  tq_transmute(select = adjusted, mutate_fun = to.period, period = "months")

monthly_gold <- tq_get(ticker_gold, from = start_date, to = end_date) |>
  drop_na(adjusted) |>
  tq_transmute(select = adjusted, mutate_fun = to.period, period = "months")

head(monthly_btc, 10)
head(monthly_gold, 10)
```

### Importing of DJIA (backup data source in case of issues with the Excel file)
```{r}
#| label: task-2-2-dj

monthly_dj <- tq_get("^DJI", from = "2015-01-01", to = "2026-02-28") |>
  drop_na(adjusted) |>
  tq_transmute(select = adjusted, mutate_fun = to.period, period = "months") |>
  rename(date = date, close_value = adjusted)
```


### Generate monthly closing prices  for all three assets and visualise them on a single faceted line plot with free y-axes to compare their price trajectories over time.
```{r}
#| label: task-2-2-2
#| fig-width: 10
#| fig-height: 6

# Standardise each asset into same 3-column format
dj_plot <- monthly_dj |>
  select(month = date, close_value) |>
  mutate(asset = "DJIA")

btc_plot <- monthly_btc |>
  select(month = date, close_value = adjusted) |>
  mutate(asset = "BTC")

gold_plot <- monthly_gold |>
  select(month = date, close_value = adjusted) |>
  mutate(asset = "Gold")

# Stack all 3 into one long data frame
combined <- bind_rows(dj_plot, btc_plot, gold_plot) |>
  mutate(asset = factor(asset, levels = c("DJIA", "BTC", "Gold")))

ggplot(combined, aes(x = month, y = close_value, color = asset)) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~asset, scales = "free_y", ncol = 1) +
  scale_color_manual(values = c("DJIA" = "steelblue", 
                                 "BTC" = "orange", 
                                 "Gold" = "gold3")) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_date(date_breaks = "2 years", date_labels = "%Y") +
  labs(
    title = "Monthly Closing Prices: DJIA, BTC, and Gold (2015–2026)",
    x = "Date",
    y = "Price (USD)",
    color = "Asset"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

### Compute the drawdown for each asset and identify the largest drawdown observed in the period for each asset.
```{r}
#| label: task-2-2-3

compute_drawdown <- function(prices) {
  drawdowns <- numeric(length(prices))
  peak <- prices[1]
  for (i in seq_along(prices)) {
    if (prices[i] > peak) peak <- prices[i]
    drawdowns[i] <- (prices[i] - peak) / peak * 100
  }
  return(drawdowns)
}

drawdowns_gold <- compute_drawdown(monthly_gold$adjusted)
drawdowns_dj   <- compute_drawdown(monthly_dj$close_value)
drawdowns_btc  <- compute_drawdown(monthly_btc$adjusted)

biggest_drawdown_gold <- min(drawdowns_gold)
biggest_drawdown_dj   <- min(drawdowns_dj)
biggest_drawdown_btc  <- min(drawdowns_btc)

biggest_drawdown_gold
biggest_drawdown_dj
biggest_drawdown_btc
```

### Results Analysis
The drawdown analysis confirms our hypothesis that volatility increases from Gold to DJIA to BTC. Gold recorded the smallest largest drawdown at approximately -17.07%, reflecting its reputation as a stable store of value over centuries. The DJIA showed a moderate drawdown of -23.20%, consistent with the broader stock market's resilience over time. BTC, however, experienced the largest drawdown at approximately -75.57%, validating its classification as a highly volatile speculative asset. These results support the conclusion that volatility increases progressively from Gold to DJIA to BTC.





---

# Task 3: Visualizing Singapore Tourists Spending 

---

# Task 4: Reflection on the Use of GenAI


Our team used a number of GenAI Models to cross check for correctness but **[Perplexity AI](https://www.perplexity.ai)** was used as our primary GenAI assistant
throughout this assessment. Unlike a standard chatbot, Perplexity combines large language
model capabilities with real-time web search, making it especially useful for up-to-date
package documentation and code debugging.

We also cross-referenced:

- [gt package documentation](https://gt.rstudio.com/) — to verify AI-suggested function arguments
- [Quarto documentation](https://quarto.org/docs/guide/) — to confirm `.qmd` syntax and YAML structure
- [tidyverse documentation](https://www.tidyverse.org/) — to validate pipe and wrangling functions

---

## 💡 How the Tool Helped Us

The tool assisted us at **every stage** of the workflow. Below are concrete examples with
the actual prompts we used.

### Parsing Match Results

We were unsure how to loop over a wide-format matrix and produce two rows per match
(one for home, one for away). We submitted this prompt:

> *"I have a CSV where rows are home teams and columns are away teams, with scores in
> home_goals-away_goals format. How do I parse this into individual team rows in R
> using tidyverse and the |> pipe?"*

Perplexity generated the nested `for` loop with `str_split()` and `case_when()` — the
core logic of our solution — in under a minute.

### GT Table Styling

The `gt` package API is extensive. Rather than reading the full documentation, we used
targeted prompts such as:

> *"How do I highlight rows where rank <= 4 in blue and rows where rank >= 18 in red
> using tab_style() and cells_body() in gt?"*

> *"How do I colour the goal_diff column green if positive and red if negative in a gt table?"*

This allowed us to apply conditional formatting confidently without trial and error.

#### 3. Debugging the YAML Error

Early on we encountered this error while testing code in the R Console:

```
Error in -title : invalid argument to unary operator
```

We pasted it directly into Perplexity:

> *"I'm getting: Error in -title : invalid argument to unary operator. I was running
> my .qmd file content in the R Console. What is wrong?"*

Perplexity immediately identified that YAML front matter cannot be executed in the R
Console — it is only valid when Quarto renders the full `.qmd` file. This saved us from
a potentially lengthy debugging session.

### Quarto Document Structure

We used the following prompt to set up the document correctly:

> *"How do I structure a Quarto .qmd document with a YAML header, prose sections,
> and R code chunks that uses |> as the pipe and outputs a single self-contained HTML
> file?"*

---

## ⏱️ Time Saved: With vs. Without GenAI

| Task | Est. Without GenAI | Actual With GenAI |
|---|---|---|
| Parsing match results | ~45 min | ~10 min |
| GT table styling | ~60 min | ~20 min |
| Debugging YAML error | ~30 min | ~2 min |
| Quarto document structure | ~20 min | ~5 min |
| **Total** | **~155 min** | **~37 min** |

---

## ⚠️ Challenges with the Tool

Despite its usefulness, we encountered real limitations:

- **Hallucinated arguments**: Perplexity once suggested a `gt` parameter that did not
  exist in the current package version. We caught this by cross-checking the
  [official gt docs](https://gt.rstudio.com/), reinforcing that AI output must always
  be verified.
- **Context drift**: In longer sessions, the tool occasionally forgot earlier constraints
  (e.g., the `|>` pipe preference), requiring us to re-specify in follow-up prompts.
- **Cannot execute R code**: Perplexity can suggest fixes but cannot run R itself —
  every suggestion had to be tested locally in RStudio, meaning debugging was still
  a back-and-forth process.
- **Over-reliance temptation**: It was tempting to accept generated code without reading
  it. We made a deliberate rule: **understand every line before using it**, so we could
  explain our own submission confidently.

---

## 📝 Partial Prompt Log

For full transparency, here is a log of some of the major prompts submitted to Perplexity AI
during this assessment:

1. > *"I have a CSV where rows are home teams and columns are away teams with scores in
   > home_goals-away_goals format. How do I parse this into individual team rows in R
   > using tidyverse and the |> pipe?"*

2. > *"How do I rank teams in R by points, then goal difference, then goals scored as
   > tiebreakers using arrange()?"*

3. > *"How do I join team abbreviations to full names using left_join() in tidyverse?"*

4. > *"How do I highlight rows where rank <= 4 in blue and rank >= 18 in red using
   > tab_style() and cells_body() in gt?"*

5. > *"How do I colour the goal_diff column green if positive and red if negative
   > in a gt table?"*

6. > *"How do I add a champion footnote to rank 1 using tab_footnote() in gt?"*

7. > *"I'm getting: Error in -title : invalid argument to unary operator when running
   > my .qmd content in the R Console. What is wrong?"*

8. > *"How do I structure a Quarto .qmd document with embed-resources: true and use
   > |> as the pipe throughout?"*

---

## 💬 Final Thoughts

GenAI compressed roughly **155 minutes of estimated work into ~37 minutes** — a 4x
productivity gain. However, this only worked because we used it as a **Collaborative Partner**, not an answer machine. The workflow that served us best was:

1. **Frame the problem clearly** — include the data structure and expected output in the prompt
2. **Iterate** — refine prompts when the first response was not precise enough
3. **Verify** — always cross-check against official documentation
4. **Understand** — read every line of generated code before using it

We believe GenAI literacy is knowing *how* to prompt effectively and *when* to question
the output which is as important a skill as the R programming itself. 

