---
title: "Week 11 Lab: Skill Wrap-Up"
subtitle: "AAI1001 – Data Engineering and Visualization | AY25/26 Tri 2 | Assessment 5 (Teamwork 2)"
date: today
date-format: "dddd MMM DD, YYYY"
author:
  - name: "Goh Jason Fadil (2502643)"
    affiliations:
      - name: "Singapore Institute of Technology"
  - name: "Venus Low (2503741)"
    affiliations:
      - name: "Singapore Institute of Technology"
  - name: "Choong Wan Qin (2402311)"
    affiliations:
      - name: "Singapore Institute of Technology"
  - name: "Chi Wei Seng (2401493)"
    affiliations:
      - name: "Singapore Institute of Technology"
  - name: "Koh Soon Chuan (2400951)"
    affiliations:
      - name: "Singapore Institute of Technology"

format:
  html:
    embed-resources: true
    number-sections: true
    toc: true
    toc-depth: 3
    toc-title: "Table of Contents"
    toc-location: left
    theme: flatly
    highlight-style: github
    code-fold: false
    df-print: paged
execute:
  echo: true
  warning: false
  message: false
---

# Task 1: English Premier League Table 2019–2020

This section of the document contains the **English Premier League 2019/20 Final Standings** from raw match results using `tidyverse` for data wrangling and `gt` for table styling.

---

## Load Libraries

```{r}
library(tidyverse)
library(readxl)
library(gt)
```

---

## Load Data

```{r}
results_raw <- read_csv("epl_results.csv")
teams_df    <- read_csv("epl_teams.csv")

teams <- teams_df$abbr
```

---

## Parse Match Results

Each match is parsed into two rows — one per team — recording goals scored, goals conceded, and points earned.

```{r}
records <- list()

for (i in seq_len(nrow(results_raw))) {
  home <- results_raw$home_team[i]

  for (away in teams) {
    if (away == home) next

    score <- results_raw[[away]][i]
    if (is.na(score) || score == "-") next

    parts <- str_split(score, "-")[[1]]
    hg    <- as.integer(parts[1])
    ag    <- as.integer(parts[2])

    # Home team row
    records[[length(records) + 1]] <- tibble(
      team          = home,
      goals_for     = hg,
      goals_against = ag,
      pts           = case_when(hg > ag ~ 3L, hg == ag ~ 1L, TRUE ~ 0L)
    )

    # Away team row
    records[[length(records) + 1]] <- tibble(
      team          = away,
      goals_for     = ag,
      goals_against = hg,
      pts           = case_when(ag > hg ~ 3L, ag == hg ~ 1L, TRUE ~ 0L)
    )
  }
}

matches <- bind_rows(records)
```

---

## Aggregate & Rank Teams

Teams are ranked by **points**, then **goal difference**, then **goals scored** as tiebreakers.

```{r}
league <- matches |>
  group_by(team) |>
  summarise(
    points        = sum(pts),
    goals_scored  = sum(goals_for),
    goals_against = sum(goals_against),
    .groups       = "drop"
  ) |>
  mutate(goal_diff = goals_scored - goals_against) |>
  arrange(desc(points), desc(goal_diff), desc(goals_scored)) |>
  mutate(rank = row_number()) |>
  left_join(teams_df |> rename(team_full = name), by = c("team" = "abbr")) |>
  select(rank, team = team_full, points, goal_diff, goals_scored)
```

---

## Generate League Table

Top 4 qualify for the **UEFA Champions League** (highlighted in blue); bottom 3 are **relegated** (highlighted in red).

```{r}
league |>
  gt() |>

  # Title & subtitle
  tab_header(
    title    = md("**\u26bd Premier League 2019/20 \u2014 Final Standings**"),
    subtitle = md("*Points \u00b7 Goal Difference \u00b7 Goals Scored*")
  ) |>

  # Column labels
  cols_label(
    rank         = md("**#**"),
    team         = md("**Club**"),
    points       = md("**Pts**"),
    goal_diff    = md("**GD**"),
    goals_scored = md("**GF**")
  ) |>

  # Alignment
  cols_align(align = "center", columns = c(rank, points, goal_diff, goals_scored)) |>
  cols_align(align = "left",   columns = team) |>

  # Row striping
  opt_row_striping() |>

  # Font
  opt_table_font(font = list(google_font("Roboto"), default_fonts())) |>

  # Global table options
  tab_options(
    heading.background.color       = "#38003c",
    heading.title.font.size        = px(20),
    heading.subtitle.font.size     = px(13),
    column_labels.background.color = "#00ff85",
    column_labels.font.weight      = "bold",
    column_labels.font.size        = px(13),
    row.striping.background_color  = "#f9f9f9",
    table.border.top.color         = "#38003c",
    table.border.bottom.color      = "#38003c",
    table_body.hlines.color        = "#e0e0e0",
    table.font.size                = px(13),
    data_row.padding               = px(6)
  ) |>

  # White text on dark purple heading
  tab_style(
    style     = list(cell_text(color = "white", weight = "bold")),
    locations = cells_title(groups = c("title", "subtitle"))
  ) |>

  # Top 4 — Champions League — blue
  tab_style(
    style     = list(cell_fill(color = "#cce5ff"), cell_text(weight = "bold")),
    locations = cells_body(rows = rank <= 4)
  ) |>

  # Bottom 3 — Relegation — red
  tab_style(
    style     = list(cell_fill(color = "#ffd6d6"), cell_text(color = "#cc0000")),
    locations = cells_body(rows = rank >= 18)
  ) |>

  # Bold rank column
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_body(columns = rank)
  ) |>

  # Positive GD in green
  tab_style(
    style     = cell_text(color = "#007a33", weight = "bold"),
    locations = cells_body(columns = goal_diff, rows = goal_diff > 0)
  ) |>

  # Negative GD in red
  tab_style(
    style     = cell_text(color = "#cc0000", weight = "bold"),
    locations = cells_body(columns = goal_diff, rows = goal_diff < 0)
  ) |>

  # Source note legend
  tab_source_note(
    source_note = md(
      "\U0001f535 Top 4: **UEFA Champions League** &nbsp;|&nbsp; \U0001f534 Bottom 3: **Relegated**"
    )
  ) |>

  # Champion footnote
  tab_footnote(
    footnote  = md("\U0001f3c6 **League Champions**"),
    locations = cells_body(columns = team, rows = rank == 1)
  )
```

---

# Task 2: Investment Market Price Analysis

## Analyzing and Visualizing DJIA Data

::: {.callout-note}
**Dataset:** Dow Jones Industrial Average (DJIA) monthly closing values from May 1896 to present, sourced from the Bloomberg DJIA Performance Report.
:::

This section covers the import, cleaning, and exploratory analysis of the Dow Jones Industrial Average (DJIA) monthly performance data sourced from Bloomberg.

### Import and Clean Data

The raw Excel file contains 5 header rows of metadata before the actual data begins. We skip these and extract only the two relevant columns — `Effective Date` and `Close Value` — coercing them to the appropriate R types.

```{r}
#| label: import-data

library(tidyverse)
library(readxl)

monthly_dj <- read_excel("dja-performance-report-monthly.xlsx", skip = 5) |>
  select(`Effective Date`, `Close Value`) |>
  rename(date = `Effective Date`, close_value = `Close Value`) |>
  mutate(
    date = as.Date(date, format = "%m/%d/%Y"),
    close_value = as.numeric(close_value)
  ) |>
  drop_na() |>
  arrange(date)

monthly_dj |> slice_head(n = 10)
```

### Identify Gap Months

We identify months where the gap to the next recorded date exceeds one month, indicating missing data in the series. The `months_to_next_row` column captures the number of calendar months between each row and the one that follows it.

```{r}
#| label: gap-months

gap_months <- monthly_dj |>
  mutate(
    month = month(date),
    year = year(date),
    months_to_next_row = interval(date, lead(date)) %/% months(1)
  ) |>
  filter(months_to_next_row > 1) |>
  select(date, close_value, month, year, months_to_next_row)

gap_months
```

::: {.callout-important}
**Historical Context:** The gap spans August to December 1914 — five missing months. This coincides with the New York Stock Exchange closure at the outbreak of World War I, as European nations liquidated US securities, threatening financial panic. The NYSE suspended trading from July 31 to December 12, 1914.
:::

### Visualizing DJIA Performance by Month

Monthly relative returns are computed and visualised using two complementary approaches to reveal distributional patterns across the calendar year. The `rel_change_pct` column captures the month-on-month percentage change in closing value, and `month_abb` provides an ordered factor for correct calendar ordering in plots.

```{r}
#| label: append-columns

monthly_dj <- monthly_dj |>
  mutate(
    rel_change_pct = (close_value / lag(close_value) - 1) * 100,
    month_abb = factor(month.abb[month(date)], levels = month.abb)
  )
```

#### Visualization A — Violin and Box Plots

The violin plot shows the full distribution of monthly returns, while the overlaid box plot highlights the interquartile range and median. The red line connects successive monthly medians to reveal any seasonal trends across the year.

```{r}
#| label: vis-a
#| fig-width: 10
#| fig-height: 6

median_line <- monthly_dj |>
  drop_na(rel_change_pct) |>
  group_by(month_abb) |>
  summarise(median_val = median(rel_change_pct))

monthly_dj |>
  drop_na(rel_change_pct) |>
  ggplot(aes(x = month_abb, y = rel_change_pct)) +
  geom_violin(fill = "steelblue", alpha = 0.4, trim = FALSE) +
  geom_boxplot(width = 0.2, outlier.shape = NA, coef = 0) +
  geom_line(
    data = median_line,
    aes(x = month_abb, y = median_val, group = 1),
    colour = "firebrick", linewidth = 0.8
  ) +
  geom_point(
    data = median_line,
    aes(x = month_abb, y = median_val),
    colour = "firebrick", size = 2
  ) +
  labs(
    title = "DJIA Monthly Returns — Violin and Box Plots",
    subtitle = "Red line connects successive monthly medians",
    caption = "Source: Bloomberg DJIA Performance Report",
    x = "Month",
    y = "Relative Change (%)"
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"))
```

#### Visualization B — Faceted Frequency Polygons

Each facet shows the distribution of monthly returns as a frequency polygon with a rug plot along the x-axis indicating individual observations. The dashed red vertical line marks the mean return for each month.

```{r}
#| label: vis-b
#| fig-width: 10
#| fig-height: 10

mean_line <- monthly_dj |>
  drop_na(rel_change_pct) |>
  group_by(month_abb) |>
  summarise(mean_val = mean(rel_change_pct))

monthly_dj |>
  drop_na(rel_change_pct) |>
  ggplot(aes(x = rel_change_pct)) +
  geom_freqpoly(binwidth = 2, colour = "steelblue") +
  geom_rug(alpha = 0.4) +
  geom_vline(
    data = mean_line,
    aes(xintercept = mean_val),
    linetype = "dashed",
    colour = "firebrick"
  ) +
  facet_wrap(~ month_abb, nrow = 4, ncol = 3) +
  labs(
    title = "DJIA Monthly Returns — Frequency Polygons by Month",
    subtitle = "Dashed red line indicates the mean return per month",
    caption = "Source: Bloomberg DJIA Performance Report",
    x = "Relative Change (%)",
    y = "Count"
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title = element_text(face = "bold"))
```

---

# Task 3: Visualizing Singapore Tourists Spending 

---

# Task 4: Reflection on the Use of GenAI

---
