---
title: "Team 30 - Assignment 2 Teamwork"
author: "Goh Jason Fadil (2502643), Venus Low (2503741), Choong Wan Qin (2402311), Chi Wei Seng (2401493), Koh Soon Chuan (2400951)"
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

# Task 1: English Premier League Table 2019–2020

This section of the document contains the **English Premier League 2019/20 Final Standings** from raw match results using `tidyverse` for data wrangling and `gt` for table styling.

---

### 1.1 Load Libraries

```{r}
library(tidyverse)
library(gt)
```

---

### 1.2 Load Data

```{r}
results_raw <- read_csv("epl_results.csv")
teams_df    <- read_csv("epl_teams.csv")

teams <- teams_df$abbr
```

---

### 1.3 Parse Match Results

Each match is parsed into two rows — one per team — recording goals scored, goals conceded, and points earned.

```{r}
records <- list()

for (i in seq_len(nrow(results_raw))) {
  home <- results_raw$home_team[i]

  for (away in teams) {
    if (away == home) next

    score <- results_raw[[away]][i]
    if (is.na(score) || score == "-") next

    parts <- str_split(score, "-")[[1]]
    hg    <- as.integer(parts[1])
    ag    <- as.integer(parts[2])

    # Home team row
    records[[length(records) + 1]] <- tibble(
      team          = home,
      goals_for     = hg,
      goals_against = ag,
      pts           = case_when(hg > ag ~ 3L, hg == ag ~ 1L, TRUE ~ 0L)
    )

    # Away team row
    records[[length(records) + 1]] <- tibble(
      team          = away,
      goals_for     = ag,
      goals_against = hg,
      pts           = case_when(ag > hg ~ 3L, ag == hg ~ 1L, TRUE ~ 0L)
    )
  }
}

matches <- bind_rows(records)
```

---

### 1.4 Aggregate & Rank Teams

Teams are ranked by **points**, then **goal difference**, then **goals scored** as tiebreakers.

```{r}
league <- matches |>
  group_by(team) |>
  summarise(
    points        = sum(pts),
    goals_scored  = sum(goals_for),
    goals_against = sum(goals_against),
    .groups       = "drop"
  ) |>
  mutate(goal_diff = goals_scored - goals_against) |>
  arrange(desc(points), desc(goal_diff), desc(goals_scored)) |>
  mutate(rank = row_number()) |>
  left_join(teams_df |> rename(team_full = name), by = c("team" = "abbr")) |>
  select(rank, team = team_full, points, goal_diff, goals_scored)
```

---

### 1.5 Generate League Table

Top 4 qualify for the **UEFA Champions League** (highlighted in blue); bottom 3 are **relegated** (highlighted in red).

```{r}
league |>
  gt() |>

  # Title & subtitle
  tab_header(
    title    = md("**\u26bd Premier League 2019/20 \u2014 Final Standings**"),
    subtitle = md("*Points \u00b7 Goal Difference \u00b7 Goals Scored*")
  ) |>

  # Column labels
  cols_label(
    rank         = md("**#**"),
    team         = md("**Club**"),
    points       = md("**Pts**"),
    goal_diff    = md("**GD**"),
    goals_scored = md("**GF**")
  ) |>

  # Alignment
  cols_align(align = "center", columns = c(rank, points, goal_diff, goals_scored)) |>
  cols_align(align = "left",   columns = team) |>

  # Row striping
  opt_row_striping() |>

  # Font
  opt_table_font(font = list(google_font("Roboto"), default_fonts())) |>

  # Global table options
  tab_options(
    heading.background.color       = "#38003c",
    heading.title.font.size        = px(20),
    heading.subtitle.font.size     = px(13),
    column_labels.background.color = "#00ff85",
    column_labels.font.weight      = "bold",
    column_labels.font.size        = px(13),
    row.striping.background_color  = "#f9f9f9",
    table.border.top.color         = "#38003c",
    table.border.bottom.color      = "#38003c",
    table_body.hlines.color        = "#e0e0e0",
    table.font.size                = px(13),
    data_row.padding               = px(6)
  ) |>

  # White text on dark purple heading
  tab_style(
    style     = list(cell_text(color = "white", weight = "bold")),
    locations = cells_title(groups = c("title", "subtitle"))
  ) |>

  # Top 4 — Champions League — blue
  tab_style(
    style     = list(cell_fill(color = "#cce5ff"), cell_text(weight = "bold")),
    locations = cells_body(rows = rank <= 4)
  ) |>

  # Bottom 3 — Relegation — red
  tab_style(
    style     = list(cell_fill(color = "#ffd6d6"), cell_text(color = "#cc0000")),
    locations = cells_body(rows = rank >= 18)
  ) |>

  # Bold rank column
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = cells_body(columns = rank)
  ) |>

  # Positive GD in green
  tab_style(
    style     = cell_text(color = "#007a33", weight = "bold"),
    locations = cells_body(columns = goal_diff, rows = goal_diff > 0)
  ) |>

  # Negative GD in red
  tab_style(
    style     = cell_text(color = "#cc0000", weight = "bold"),
    locations = cells_body(columns = goal_diff, rows = goal_diff < 0)
  ) |>

  # Source note legend
  tab_source_note(
    source_note = md(
      "\U0001f535 Top 4: **UEFA Champions League** &nbsp;|&nbsp; \U0001f534 Bottom 3: **Relegated**"
    )
  ) |>

  # Champion footnote
  tab_footnote(
    footnote  = md("\U0001f3c6 **League Champions**"),
    locations = cells_body(columns = team, rows = rank == 1)
  )
```

---

# Task 2: Investment Market Price Analysis

---

# Task 3: Visualizing Singapore Tourists Spending 

---

# Task 4: Reflection on the Use of GenAI

###🤖 GenAI Tool Used

Our team used **[Perplexity AI](https://www.perplexity.ai)** as our primary GenAI assistant
throughout this assessment. Unlike a standard chatbot, Perplexity combines large language
model capabilities with real-time web search, making it especially useful for up-to-date
package documentation and code debugging.

We also cross-referenced:

- [gt package documentation](https://gt.rstudio.com/) — to verify AI-suggested function arguments
- [Quarto documentation](https://quarto.org/docs/guide/) — to confirm `.qmd` syntax and YAML structure
- [tidyverse documentation](https://www.tidyverse.org/) — to validate pipe and wrangling functions

---

### 💡 How the Tool Helped Us

The tool assisted us at **every stage** of the workflow. Below are concrete examples with
the actual prompts we used.

#### 1. Parsing Match Results

We were unsure how to loop over a wide-format matrix and produce two rows per match
(one for home, one for away). We submitted this prompt:

> *"I have a CSV where rows are home teams and columns are away teams, with scores in
> home_goals-away_goals format. How do I parse this into individual team rows in R
> using tidyverse and the |> pipe?"*

Perplexity generated the nested `for` loop with `str_split()` and `case_when()` — the
core logic of our solution — in under a minute.

#### 2. GT Table Styling

The `gt` package API is extensive. Rather than reading the full documentation, we used
targeted prompts such as:

> *"How do I highlight rows where rank <= 4 in blue and rows where rank >= 18 in red
> using tab_style() and cells_body() in gt?"*

> *"How do I colour the goal_diff column green if positive and red if negative in a gt table?"*

This allowed us to apply conditional formatting confidently without trial and error.

#### 3. Debugging the YAML Error

Early on we encountered this error while testing code in the R Console:

```
Error in -title : invalid argument to unary operator
```

We pasted it directly into Perplexity:

> *"I'm getting: Error in -title : invalid argument to unary operator. I was running
> my .qmd file content in the R Console. What is wrong?"*

Perplexity immediately identified that YAML front matter cannot be executed in the R
Console — it is only valid when Quarto renders the full `.qmd` file. This saved us from
a potentially lengthy debugging session.

#### 4. Quarto Document Structure

We used the following prompt to set up the document correctly:

> *"How do I structure a Quarto .qmd document with a YAML header, prose sections,
> and R code chunks that uses |> as the pipe and outputs a single self-contained HTML
> file?"*

---

### ⏱️ Time Saved: With vs. Without GenAI

| Task | Est. Without GenAI | Actual With GenAI |
|---|---|---|
| Parsing match results | ~45 min | ~10 min |
| GT table styling | ~60 min | ~20 min |
| Debugging YAML error | ~30 min | ~2 min |
| Quarto document structure | ~20 min | ~5 min |
| **Total** | **~155 min** | **~37 min** |

---

### ⚠️ Challenges with the Tool

Despite its usefulness, we encountered real limitations:

- **Hallucinated arguments**: Perplexity once suggested a `gt` parameter that did not
  exist in the current package version. We caught this by cross-checking the
  [official gt docs](https://gt.rstudio.com/), reinforcing that AI output must always
  be verified.
- **Context drift**: In longer sessions, the tool occasionally forgot earlier constraints
  (e.g., the `|>` pipe preference), requiring us to re-specify in follow-up prompts.
- **Cannot execute R code**: Perplexity can suggest fixes but cannot run R itself —
  every suggestion had to be tested locally in RStudio, meaning debugging was still
  a back-and-forth process.
- **Over-reliance temptation**: It was tempting to accept generated code without reading
  it. We made a deliberate rule: **understand every line before using it**, so we could
  explain our own submission confidently.

---

### 📝 Full Prompt Log

For full transparency, here is a log of all major prompts submitted to Perplexity AI
during this assessment:

1. > *"I have a CSV where rows are home teams and columns are away teams with scores in
   > home_goals-away_goals format. How do I parse this into individual team rows in R
   > using tidyverse and the |> pipe?"*

2. > *"How do I rank teams in R by points, then goal difference, then goals scored as
   > tiebreakers using arrange()?"*

3. > *"How do I join team abbreviations to full names using left_join() in tidyverse?"*

4. > *"How do I highlight rows where rank <= 4 in blue and rank >= 18 in red using
   > tab_style() and cells_body() in gt?"*

5. > *"How do I colour the goal_diff column green if positive and red if negative
   > in a gt table?"*

6. > *"How do I add a champion footnote to rank 1 using tab_footnote() in gt?"*

7. > *"I'm getting: Error in -title : invalid argument to unary operator when running
   > my .qmd content in the R Console. What is wrong?"*

8. > *"How do I structure a Quarto .qmd document with embed-resources: true and use
   > |> as the pipe throughout?"*

---

### 💬 Final Thoughts

GenAI compressed roughly **155 minutes of estimated work into ~37 minutes** — a 4x
productivity gain. However, this only worked because we used it as a **Collaborative Partner**, not an answer machine. The workflow that served us best was:

1. **Frame the problem clearly** — include the data structure and expected output in the prompt
2. **Iterate** — refine prompts when the first response was not precise enough
3. **Verify** — always cross-check against official documentation
4. **Understand** — read every line of generated code before using it

We believe GenAI literacy is knowing *how* to prompt effectively and *when* to question
the output which is as important a skill as the R programming itself. 
